Enforcing Well-bracketed Control Flow and Stack Encapsulation using Linear Capabilities

### 2018-01-12 presentations/prisc18-01-2017-linear-caps/presentation.tex
Small fix in the PriSC presentation.

@@

  \vspace{-1cm}
  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
    % \draw<1-6> (-5,0) node {};
    % \draw (5,0) node {};
    % \draw (0,-14) node {};
    \draw (0,0) node[draw=black] (ir) {...};
    \draw node[draw=black,rounded corners,above=of ir] (source) {Source Language};
    \draw (source) edge[bend left, ->] (ir);
    \begin{onlyenv}<1-3>
      \draw node[draw=black,rounded corners,below=of ir] (target) {Target Language};
      \draw (ir) edge[bend left, ->] (target) ;
    \end{onlyenv}
    \begin{onlyenv}<2>
      \draw node[right=1cm of ir, line width=1mm, font=\fontfamily{ppl}\fontsize{1cm}{1.5cm}\selectfont] {?};
    \end{onlyenv}
    \draw<3> node[right=1cm of target, line width=1mm, font=\fontfamily{ppl}\fontsize{1.5cm}{2cm}\selectfont] {?};
    \begin{visibleenv}<4-5>
      \draw (200:5) node[draw=black,rounded corners, text width=3.5cm] (commhar) {Commodity Hardware\\(Virtual Memory)};
      \draw (ir) edge[bend right=10, ->] (commhar) ;
      \draw (230:5) node[draw=black,rounded corners, text width=3.5cm] (wasm) {WebAssembly};
      \draw (ir) edge[bend left, ->] (wasm) ;
    \end{visibleenv}
    \begin{visibleenv}<4-6>
      \draw (280:5.8) node[draw=black,rounded corners, text width=3.5cm] (dtal) {Dependently-typed\\Assembly Language};
      \draw (ir) edge[bend left=20, ->] (dtal) ;
      \draw (335:8) node[draw=black,rounded corners, text width=3.5cm] (mph) {Micro-Policies Hardware};
      \draw (ir) edge[bend right=20, ->] (mph) ;
      \draw (350:7) node[draw=black,rounded corners, text width=3.5cm] (capmach) {Capability Machines};
      \draw (ir) edge[bend left=10, ->] (capmach) ;
    \end{visibleenv}
    \begin{visibleenv}<5>
      \draw node[below left=0cm of commhar, font=\fontfamily{ppl}\fontsize{1cm}{1.5cm}\selectfont] {?};
      \draw node[below left=0cm of wasm, font=\fontfamily{ppl}\fontsize{1cm}{1.5cm}\selectfont] {?};
    \end{visibleenv}
    \begin{visibleenv}<7>
      \draw node[draw=black,rounded corners,below=of ir] (capmach2) {Capability machines};
      \draw (ir) edge[bend left, ->] (capmach2) ;
    \end{visibleenv}
    \begin{visibleenv}<8->
      \draw node[draw=black,below=of ir] (ir2) {...};
      \draw (ir) edge[bend left, ->] (ir2) ;
      \draw node[draw=black,rounded corners,below=of ir2] (capmach3) {Capability machines};
      \draw (ir2) edge[bend left, ->] (capmach3) ;
    \end{visibleenv}
    \begin{visibleenv}<9>
      \draw node[rectangle,fit=(ir2) (capmach3),draw=red] (laststeps) {};
      \draw node[above right=0cm of laststeps,cloud,red] {Let's take a look!};
    \end{visibleenv}
  \end{tikzpicture}
  \only<5>{Fine-grained encapsulation?}
  \only<8>{Well-bracketed control flow, stack encapsulation?}
\end{frame}

@@

\begin{frame}<1>[label=sarps]
  \frametitle{Stack and return pointer security}
  \begin{itemize}
  \item Desirable properties:
    \begin{itemize}
    \item Well-bracketed control flow
    \item Stack frame encapsulation
    \end{itemize}
  \item How to enforce?
    \begin{itemize}
    \item Capabilities?
      \begin{itemize}
      \item<2-> Not enough: two attacks!
      \end{itemize}
    \item<3-> CHERI Local capabilities?
      \begin{itemize}
      \item Cannot leave registers
      \item<4-> Works, but... [Skorstengaard et al., ESOP 2018]
      \end{itemize}
    \item<5-> CHERI Linear capabilities?
      \begin{itemize}
      \item Non-copyable capabilities
      \item<6> Works perfectly?
      \end{itemize}
    \end{itemize}
  \end{itemize}
\end{frame}

@@

  \begin{itemize}
  \item<2-> Formulate correctness?
    \begin{itemize}
    % \item<3-> (CFI is too weak (right flows \emph{at the right time}))
    \item<3-> Idea: well-bracketed version of target assembly
    \item<3-> Compiler replaces call/returns with calling convention
    \item<3-> Fully abstract compilation
    \end{itemize}
  \item Prove correctness?
    \begin{itemize}
    \item<4-> Back-translation as embedding
    \item<4-> Cross-language step-indexed Kripke logical relation
    \end{itemize}
  \end{itemize}
\end{frame}

>>>>

\tikzset{
    gray on/.style={alt=#1{gray}{}},
    alt/.code args={<#1>#2#3}{%
      \alt<#1>{\pgfkeysalso{#2}}{\pgfkeysalso{#3}} % \pgfkeysalso doesn't change the path
    },
}

@@

  \vspace{-1cm}
  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
    % \draw<1-6> (-5,0) node {};
    % \draw (5,0) node {};
    % \draw (0,-14) node {};
    \draw (0,0) node[draw=black] (ir) {...};
    \draw node[draw=black,rounded corners,above=of ir] (source) {Source Language};
    \draw (source) edge[bend left, ->] (ir);
    \begin{onlyenv}<1-3>
      \draw node[draw=black,rounded corners,below=of ir] (target) {Target Language};
      \draw (ir) edge[bend left, ->] (target) ;
    \end{onlyenv}
    \begin{onlyenv}<2>
      \draw node[right=1cm of ir, line width=1mm, font=\fontfamily{ppl}\fontsize{1cm}{1.5cm}\selectfont] {?};
    \end{onlyenv}
    \draw<3> node[right=1cm of target, line width=1mm, font=\fontfamily{ppl}\fontsize{1.5cm}{2cm}\selectfont] {?};
    \begin{visibleenv}<4-5>
      \draw (200:5) node[draw=black,rounded corners, text width=3.5cm] (commhar) {Commodity Hardware\\(Virtual Memory)};
      \draw (ir) edge[bend right=10, ->] (commhar) ;
      \draw (230:5) node[draw=black,rounded corners, text width=3.5cm] (wasm) {WebAssembly};
      \draw (ir) edge[bend left, ->] (wasm) ;
    \end{visibleenv}
    \begin{visibleenv}<4-7>
      \draw (280:5.8) node[gray on=<7>,draw=black,rounded corners, text width=3.5cm] (dtal) {Dependently-typed\\Assembly Language};
      \draw (ir) edge[gray on=<7>,bend left=20, ->] (dtal) ;
      \draw (335:8) node[gray on=<7>,draw=black,rounded corners, text width=3.5cm] (mph) {Micro-Policies Hardware};
      \draw (ir) edge[gray on=<7>,bend right=20, ->] (mph) ;
      \draw (350:7) node[draw=black,rounded corners, text width=3.5cm] (capmach) {Capability Machines};
      \draw (ir) edge[bend left=10, ->] (capmach) ;
    \end{visibleenv}
    \begin{visibleenv}<5>
      \draw node[below left=0cm of commhar, font=\fontfamily{ppl}\fontsize{1cm}{1.5cm}\selectfont] {?};
      \draw node[below left=0cm of wasm, font=\fontfamily{ppl}\fontsize{1cm}{1.5cm}\selectfont] {?};
    \end{visibleenv}
    \begin{visibleenv}<8>
      \draw node[draw=black,rounded corners,below=of ir] (capmach2) {Capability machines};
      \draw (ir) edge[bend left, ->] (capmach2) ;
    \end{visibleenv}
    \begin{visibleenv}<9->
      \draw node[draw=black,below=of ir] (ir2) {...};
      \draw (ir) edge[bend left, ->] (ir2) ;
      \draw node[draw=black,rounded corners,below=of ir2] (capmach3) {Capability machines};
      \draw (ir2) edge[bend left, ->] (capmach3) ;
    \end{visibleenv}
    \begin{visibleenv}<10>
      \draw node[rectangle,fit=(ir2) (capmach3),draw=red] (laststeps) {};
      \draw node[above right=0cm of laststeps,cloud,red] {Let's take a look!};
    \end{visibleenv}
  \end{tikzpicture}
  \only<5>{Fine-grained encapsulation?}
  \only<9>{Well-bracketed control flow, stack encapsulation?}
\end{frame}

@@

\begin{frame}<1>[label=sarps]
  \frametitle{Stack and return pointer security}
  \begin{itemize}
  \item Desirable properties:
    \begin{itemize}
    \item Well-bracketed control flow
    \item Stack frame encapsulation
    \end{itemize}
  \item How to enforce?
    \begin{itemize}
    \item Capabilities?
      \begin{itemize}
      \item<2-> Not enough: three attacks!
      \end{itemize}
    \item<3-> CHERI Local capabilities?
      \begin{itemize}
      \item Cannot leave registers
      \item<4-> Works, but... [Skorstengaard et al., ESOP 2018]
      \end{itemize}
    \item<5-> CHERI Linear capabilities?
      \begin{itemize}
      \item Non-copyable capabilities
      \item<6> Works perfectly?
      \end{itemize}
    \end{itemize}
  \end{itemize}
\end{frame}

@@

  \begin{itemize}
  \item<2-> Formulate correctness?
    \begin{itemize}
    % \item<3-> (CFI is too weak (right flows \emph{at the right time}))
    \item<4-> Idea: well-bracketed version of target assembly
    \item<4-> Compiler replaces call/returns with calling convention
    \item<3-> Fully abstract compilation
    \end{itemize}
  \item Prove correctness?
    \begin{itemize}
    \item<5-> Back-translation as embedding
    \item<5-> Cross-language step-indexed Kripke logical relation
    \end{itemize}
  \end{itemize}
\end{frame}
